
---
title: "Neighborhood Disadvantage and Crime Variation Over Time"
author: "Olivia Delos"
---

# Introduction
Crime rates have decreased for all types of crime since the 1990s, although most considerably for violent crimes (Papachristos 2013). However, crime rates vary depending on the social composition of a neighborhood. Disadvantaged neighborhoods are disproportionately low-income with weak social ties and high residential instability (Zuberi 2012). The concentration of poverty, violence, and high policing in disadvantaged neighborhoods hinder residents’ ability to form social cohesion and community organization, increasing residents’ risk of victimization and limiting social control (Van Wilsem, Wittebrood, De Graaf 2006). Disadvantaged neighborhoods are also often policed at higher rates than their counterparts. The high poverty and policing combined with lack of social control and low social cohesion in disadvantaged neighborhoods leads to increased crime rates in these areas (Van Wilsem et al. 2006). Violent crime, in particular, is more prevalent in urban, disadvantaged neighborhoods (Friedson and Sharkey 2015). However, few studies have examined whether the type of crimes committed in disadvantaged neighborhoods are different from those committed in less disadvantaged areas. By examining census data and crime data from Buffalo, New York I want to answer three questions: 1) Are the types of crime being committed in low-income census tracts different from those in higher income tracts? 2) Have census tracts witnessed an overall decrease in crime between 2010 and 2017? 3) If so, does the decrease differ by crime type? To answer my questions, I will compare crime rates by type of crime incident between 2010 to crime rates in 2017. I measure neighborhood disadvantage using census tract data in Buffalo, New York on median income from the 2010 and 2017 ACS. Crime data comes from the public PenData Buffalo and can be broken into five categories: assault, breaking and entering, homicide, sexual offenses, and theft. I predict that, although crime rates for all types are likely higher in lower income neighborhoods, the difference is likely greater for violent crimes. I also hypothesize that there will be a decrease in crimes rates between 2010 and 2017 for all tracts, but more noticeably for violent crimes, as opposed to property crimes.


# Materials and methods

1. Load required packages
2. Download crime data from buffalony.gov and 2010 and 2017 ACS estimates for income data in Erie County
3. Clean and join data
4. Run regression on types of crime and income level
5. Create figures

## 1. Load required packages.
```{r, message=F, warning=F}
rm(list=ls())
library(tidycensus)
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(seg)
library(sf)
library(tigris)
library(gridExtra)
library(spgwr)
```

## 2. Download the data

```{r, results="hide", cache = TRUE, message=F, warning=F}
#Load crime data
crimesf=read_sf("https://data.buffalony.gov/api/geospatial/d6g9-xbgu?method=export&format=GeoJSON")

#Load 2017 census data
v17 <- load_variables(year = 2017, dataset = "acs5", cache = TRUE)

med_income2017 <- get_acs(geography = "tract",
                        variables = c(medianinc = "B06011_001"),
                       state = "NY",
                       county = "Erie",
                       year = 2017,
                       geometry = TRUE)

totalpop2017 <- get_acs(geography = "tract",
                        variables = c(totalpop = "B01003_001"),
                       state = "NY",
                       county = "Erie",
                       year = 2017,
                       geometry = TRUE)


# Load 2010 census data
v10 <- load_variables(year = 2010, dataset = "acs5", cache = TRUE)

med_income2010 <- get_acs(geography = "tract",
                        variables = "B19013_001",
                       state = "NY",
                       county = "Erie",
                       year = 2010,
                       geometry = TRUE)

totalpop2010 <- get_acs(geography = "tract",
                        variables = c(totalpop = "B01003_001"),
                       state = "NY",
                       county = "Erie",
                       year = 2010,
                       geometry = TRUE)


```
Crop census data to Buffalo city limits
``` {r, results = "hide", cache = TRUE, message=F, warning=F}

#Crop census data to city limits
income_crop2017 <- med_income2017[c(1:70, 228, 229, 221:227),]
income_crop2010 <- med_income2010[c(1:70, 228, 229, 221:227),]
totalpop17 <- totalpop2017[c(1:70, 228, 229, 221:227),]
totalpop10 <- totalpop2010[c(1:70, 228, 229, 221:227),]

#Transform crime geometry to match census tract
crimetrans <- st_transform(crimesf, st_crs(income_crop2017))

inc_tracts <- st_join(crimetrans, select(med_income2017,NAME), st_intersects)
```

## 3. Clean crime data
```{r, cache = TRUE,results="hide"}
#extract year from date
crime_date <- ymd_hms(inc_tracts$incident_datetime) %>%
    year()
  
crime_year <- cbind(crime_date, inc_tracts, stringsAsFactors = FALSE)

#Recategorize crime types
crime_year$parent_incident_type <- as.factor(crime_year$parent_incident_type) %>%
  fct_collapse(crime_year$parent_incident_type, 
                  Theft = c("Theft of Vehicle", "Theft", "Robbery"),
                  Sexual_Offenses = c("Sexual Assault", "Other Sexual Offense"),
                  Breaking_entering = "Breaking & Entering",
                  Assault = "Assault",
                  Homicide = "Homicide")

#clean/filter crime data, drop geometry
crime2017 <- crime_year %>%
     filter(crime_date == 2017) %>%
     group_by(NAME, parent_incident_type) %>%
    summarize(total = n()) %>%
    st_set_geometry(NULL)


c2017_spread <- spread(crime2017, parent_incident_type, total)

c2017_spread[is.na(c2017_spread)] <- 0

#Do the same for 2010
crime2010 <- crime_year %>%
     filter(crime_date == 2010) %>%
     group_by(NAME, parent_incident_type) %>%
    summarize(total = n()) %>%
    st_set_geometry(NULL)

c2010_spread <- spread(crime2010, parent_incident_type, total)

c2010_spread[is.na(c2010_spread)] <- 0

```

Clean income data and join datasets
```{r, cache = TRUE}
#Remove and rename columns
income_clean17 <- income_crop2017 %>%
  select(-moe, -variable, -GEOID)

income_clean10 <- income_crop2010 %>%
  select(-moe, -variable, -GEOID)

pop2017_clean <- totalpop17 %>%
  select(-moe, -variable, -GEOID)
  
pop2010_clean <- totalpop10 %>%
  select(-moe, -variable, -GEOID)

colnames(income_clean10)[2] <- "MedianInc"
colnames(income_clean17)[2] <- "MedianInc"
colnames(pop2017_clean)[2] <- "Population"
colnames(pop2010_clean)[2] <- "Population"

#Join income and crime spread
crime_income2017 <- left_join(income_clean17, c2017_spread, by = c("NAME" = "NAME"))
crime_income2010 <- left_join(income_clean10, c2010_spread, by = c("NAME" = "NAME"))

#Join income and population
inc_pop2017 <- cbind(income_clean17, pop2017_clean) %>%
  select(-NAME.1, -geometry.1)

inc_pop2010 <- cbind(income_clean10, pop2010_clean) %>%
  select(-NAME.1, -geometry.1)
```

Calculate avg crime rate per person
```{r,cache=TRUE,message=F,warning=F}
#Find mean total crime per tract
crimeAvg2017 <- crime_income2017 %>%
  group_by(NAME, MedianInc) %>%
  summarize(meancrime = mean(c(Assault, Breaking_entering, Sexual_Offenses, Theft)))

crimeAvg2010 <- crime_income2010 %>%
  group_by(NAME, MedianInc) %>%
  summarize(meancrime = mean(c(Assault, Breaking_entering, Sexual_Offenses, Theft)))

#Calculate crime rate per population
cpp17 <- crimeAvg2017[ order(match(crimeAvg2017$NAME, inc_pop2017$NAME)),]
cpp10 <- inc_pop2010[ order(match(inc_pop2010$NAME, crimeAvg2010$NAME)),]
  
crime_pop2017 <- cbind(inc_pop2017, crimeAvg2017) %>%
  select(-MedianInc.1, -geometry.1, -NAME.1)

crime_pop2010 <- cbind(inc_pop2017, crimeAvg2010) %>%
  select(-MedianInc.1, -geometry.1, -NAME.1)


cpp2017 <- crime_pop2017 %>%
  mutate(AvgCPC = (meancrime/Population))

cpp2010 <- crime_pop2010 %>%
  mutate(AvgCPC = (meancrime/Population))
```

Join population data with crime data
```{r, message=F, warning=F}
#Join population and crime_spread 2010
crimetype_pop10 <- cbind(crime_income2010, pop2010_clean)

avghom2010 <- crimetype_pop10 %>%
  mutate(AvgHom = (Homicide/Population))

avgAss2010 <- crimetype_pop10 %>%
  mutate(AvgAssault = (Assault/Population))

avgBE2010 <- crimetype_pop10 %>%
  mutate(AvgBreaking_entering = (Breaking_entering/Population))

avgTheft2010 <- crimetype_pop10 %>%
  mutate(AvgTheft = (Theft/Population))

avgSO2010 <- crimetype_pop10 %>%
  mutate(avgSO = (Sexual_Offenses/Population))

#Join population and crime_spread 2017
crimetype_pop17 <- cbind(crime_income2017, pop2017_clean)

avghom2017 <- crimetype_pop17 %>%
  mutate(AvgHom = (Homicide/Population))

avgAss2017 <- crimetype_pop17 %>%
  mutate(AvgAssault = (Assault/Population))

avgBE2017 <- crimetype_pop17 %>%
  mutate(AvgBreaking_entering = (Breaking_entering/Population))

avgTheft2017 <- crimetype_pop17 %>%
  mutate(AvgTheft = (Theft/Population))

avgSO2017 <- crimetype_pop17 %>%
  mutate(avgSO = (Sexual_Offenses/Population))
```


Clean crime data for time series analysis
```{r}
# Clean for all of Buffalo
crimeALL <- crime_year %>%
     filter(crime_date >= 2010 & crime_date <= 2017) %>%
     group_by(NAME, parent_incident_type, crime_date) %>%
    summarize(total = n()) %>%
    st_set_geometry(NULL)

# Don't group by tract name for overall time series
crimeALLplot <- crime_year %>%
  filter(crime_date >= 2010 & crime_date <= 2017) %>%
     group_by(parent_incident_type, crime_date) %>%
    summarize(total = n()) %>%
    st_set_geometry(NULL)
  
# crop crimeALL for Elmwood tracts
elmwoodcrop <- crimeALL[c(569:600,2301:2571),]

elmwood <- elmwoodcrop %>%
  group_by(parent_incident_type, crime_date) %>%
  spread(parent_incident_type, total)

elmwood[is.na(elmwood)] <- 0

elmwoodavg <- summarize(elmwood, avg = mean(c(Assault, Breaking_entering, Sexual_Offenses, Homicide, Theft)))


# Crop for waterfront tracts 
waterfrontcrop <- crimeALL[c(2263:2300,2572:2642,2675:2810,574:710),]

# Clean for waterfront
waterfront <- waterfrontcrop %>%
  group_by(parent_incident_type, crime_date) %>%
  spread(parent_incident_type, total)

waterfront[is.na(waterfront)] <- 0

waterfrontavg <- summarize(waterfront, avg = mean(c(Assault, Breaking_entering, Sexual_Offenses, Homicide, Theft)))

```
## 4. Run regression/plot the relationship between crime and median income
Plot relationship between crime and income
```{r, results='asis', cache = TRUE, message=F, warning=F}
#2010 plot
ggplot(cpp2010) + geom_point(aes(x = MedianInc, y = AvgCPC)) +
  labs(x = "Median Income 2010", y = "Average Crime Per Capita 2010", title = "Relationship between Median Income and Crime Per Capita in 2010")

#2017 plot
ggplot(cpp2017) + geom_point(aes(x = MedianInc, y = AvgCPC)) +
  labs(x = "Median Income 2017", y = "Average Crime Per Capita 2017", title = "Relationship between Median Income and Crime Per Capita in 2017")

```

Relationship between crime and income 2010
```{r}
#2010 regression
Assault2010_reg <- lm(Assault~MedianInc, data = crimetype_pop10)
summary(Assault2010_reg)

Theft2010_reg <- lm(Theft~MedianInc, data = crimetype_pop10)
summary(Theft2010_reg)

BE2010_reg <- lm(Breaking_entering~MedianInc, data = crimetype_pop10)
summary(BE2010_reg)

#No correlation
SO2010_reg <- lm(Sexual_Offenses~MedianInc, data = crimetype_pop10)
summary(SO2010_reg)

#Highest correlation
Hom2010_reg <- lm(Homicide~MedianInc, data = crimetype_pop10)
summary(Hom2010_reg)

```

Relationship between crime and income 2017
```{r}
#2017 regression
Assault2017_reg <- lm(Assault~MedianInc, data = crimetype_pop17)
summary(Assault2017_reg)

Theft2017_reg <- lm(Theft~MedianInc, data = crimetype_pop17)
summary(Theft2017_reg)

BE2017_reg <- lm(Breaking_entering~MedianInc, data = crimetype_pop17)
summary(BE2017_reg)

#No correlation
SO2017_reg <- lm(Sexual_Offenses~MedianInc, data = crimetype_pop17)
summary(SO2017_reg)

#Highest correlation
Hom2017_reg <- lm(Homicide~MedianInc, data = crimetype_pop17)
summary(Hom2017_reg)
```

# Results


### Plot median income and crime per capita for 2010 and 2017

This figure is a map of overall crime frequency over time by type
```{r, fig.width=10, fig.height=5, fig.cap="Crime Over Time"}
ggplot(crimeALLplot) + geom_line(aes(x = crime_date, y = total, color = parent_incident_type)) + 
  labs(x = "Year", y = "Total crime", color = "Crime type")
```
This figure is a map of the average crime rate for all types betweeen 2010 and 2017 in Elmwood Village
```{r, fig.width=10, fig.height=5, fig.cap="Average crime between 2010 and 2017 in Elmwood Village"}
ggplot(elmwoodavg) + geom_line(aes(x = crime_date, y = avg))
```
This figure is a map of average crime rate for all types between 2010 and 2017 in Buffalo's Waterfront Area.
```{r, fig.width=10, fig.height=5, fig.cap="Average crime between 2010 and 2017 in Buffalo's Waterfront Area"}
ggplot(waterfrontavg) + geom_line(aes(x = crime_date, y = avg))
```

This figure is a map of median income and average crime per capita for 2010 by census tract in Buffalo, NY
```{r, fig.width=10, fig.height=5, fig.cap="2010 Income and Crime"}

mapinc10 <- ggplot(cpp2010) + geom_sf(aes(fill = MedianInc)) +
  scale_fill_viridis_c() + labs(fill = "Median Income") +
  theme(panel.background = element_blank())

mapcrime10 <- ggplot(cpp2010) + geom_sf(aes(fill = AvgCPC)) + 
  scale_fill_viridis_c() + labs(fill = "Average CPC") +
  theme(panel.background = element_blank())

grid.arrange(mapinc10, mapcrime10, nrow = 1)

```

This figure is a map of median income and average crime per capita for 2017 by census tract in Buffalo, NY
```{r, fig.width=10, fig.height=5, fig.cap="2017 Income and Crime"}
mapinc17 <- ggplot(cpp2017) + geom_sf(aes(fill = MedianInc)) +
  scale_fill_viridis_c() + labs(fill = "Median Income") + 
  theme(panel.background = element_blank())

mapcrime17 <- ggplot(cpp2017) + geom_sf(aes(fill = AvgCPC)) + 
  scale_fill_viridis_c() + labs(fill = "Average CPC") +
  theme(panel.background = element_blank())

grid.arrange(mapinc17, mapcrime17, nrow = 1)
```

## Plot crime by type

Homicide 2010
```{r,fig.width=10, fig.height=5}

hom2010map <- ggplot(avghom2010) + geom_sf(aes(fill = AvgHom)) + scale_fill_viridis_c() + 
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Homicide PC")

grid.arrange(mapinc10, hom2010map, nrow = 1)
```

Homicide 2017
```{r,fig.width=10, fig.height=5}

hom2017map <- ggplot(avghom2017) + geom_sf(aes(fill = AvgHom)) + scale_fill_viridis_c() +
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Homicide PC")

grid.arrange(mapinc17, hom2017map, nrow = 1)
```

Assault2010
```{r,fig.width=10, fig.height=5}
assault2010map <- ggplot(avgAss2010) + geom_sf(aes(fill = AvgAssault)) + scale_fill_viridis_c() + 
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Assault PC")

grid.arrange(mapinc10, assault2010map, nrow = 1)

```
Assault2017
```{r,fig.width=10, fig.height=5}
assault2017map <- ggplot(avgAss2017) + geom_sf(aes(fill = AvgAssault)) + scale_fill_viridis_c() +
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Assault PC")

grid.arrange(mapinc17, assault2017map, nrow = 1)
```

Breaking & Entering 2010
```{r,fig.width=10, fig.height=5}
be2010map <- ggplot(avgBE2010) + geom_sf(aes(fill = AvgBreaking_entering)) + scale_fill_viridis_c() + 
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Break&Ent. PC")

grid.arrange(mapinc10, be2010map, nrow = 1)
```

Breaking & Entering 2017
```{r,fig.width=10, fig.height=5}
be2017map <- ggplot(avgBE2017) + geom_sf(aes(fill = AvgBreaking_entering)) + scale_fill_viridis_c() + 
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Break&Ent. PC")

grid.arrange(mapinc17, be2017map, nrow = 1)
```
Theft 2010
```{r,fig.width=10, fig.height=5}
theft2010map <- ggplot(avgTheft2010) + geom_sf(aes(fill = AvgTheft)) + scale_fill_viridis_c() + 
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Theft PC")

grid.arrange(mapinc10, theft2010map, nrow = 1)
```
Theft 2017
```{r,fig.width=10, fig.height=5}
theft2017map <- ggplot(avgTheft2017) + geom_sf(aes(fill = AvgTheft)) + scale_fill_viridis_c() + 
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Theft PC")

grid.arrange(mapinc17, theft2017map, nrow = 1)
```
Sexual_offenses 2010
```{r,fig.width=10, fig.height=5}
so2010map <- ggplot(avgSO2017) + geom_sf(aes(fill = avgSO)) + scale_fill_viridis_c() + 
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Sex. Off. PC")

grid.arrange(mapinc10, so2010map, nrow = 1)
```

Sexual_offenses 2017
```{r,fig.width=10, fig.height=5}
so2017map <- ggplot(avgSO2017) + geom_sf(aes(fill = avgSO)) + scale_fill_viridis_c() + 
  theme(panel.background = element_blank()) +
  labs(fill = "Avg Sex. Off. PC")

grid.arrange(mapinc17, so2017map, nrow = 1)
```
# Conclusions
Overall, it seems that median income has a moderate impact on crime. Crime has decreased overall in Buffalo, but there is no pattern in decrease by any particular crime type. The relationship between crime and median income is higher when examining violent crimes.......NAME CRIMES specifically show a higher rate of crime for WHICH TYPES in tracts that have lower median incomes....These results suggest that median income plays a role in crime rate, but is not the sole determinant. There are likely other factors that determine crime rates in an area, and these may differ depending on the area being examined.

# References

https://data.buffalony.gov/Public-Safety/Crime-Incidents/d6g9-xbgu

Friedson, Michael and Patrick Sharkey. 2015. "Violence and Neighborhood Disadvantage after the Crime Decline." *The American Academy of Political & Social Science* 660. 

Papachristos, Andrew V. 2013. "48 Years of Crime in Chicago: A Descriptive Analysis of Serious Crime Trends from 1965 to 2013." *Institution for Social and Policy Studies*. 

Van Wilsem, Johan. Karin Whitebrood, Nan Dirk De Graaf. 2006. "Socioeconomic Dynamics of Neighborhoods and the Risk of Crime Victimization: A Multilevel Study of Improving, Declining, and Stable Areas in the Netherlands." *Social Problems* 53(2):226-247.

Zuberi, Anita. 2012. “Neighborhood Poverty and Children’s Exposure to Danger: Examining Gender Differences in Impacts of the Moving to Opportunity Experiment.” *Social Science Research* 41:788-801.